---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geomander

<!-- badges: start -->
![GitHub](https://img.shields.io/github/license/christopherkenny/geomander)
[![R-CMD-check](https://github.com/christopherkenny/geomander/workflows/R-CMD-check/badge.svg)](https://github.com/christopherkenny/geomander/actions)
<!-- badges: end -->

Focuses on creating data sets and other tools that help make understanding gerrymandering faster and easier. Designed for easy preparation to run simulation analysis with the R package redist, but is aimed at the geographic aspects of redistricting, not partitioning methods. Most of these tools are gathered from seminar papers and do not correspond to a single publication.

## Installation
<!---
You can install the released version of geomander from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("geomander")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("christopherkenny/geomander")
```
--->

Currently, geomander is only available from [GitHub](https://github.com/christopherkenny/geomander) and can be installed with:
```r
# install.packages('remotes')
remotes::install_github('christopherkenny/geomander')
```

Needs redist back on CRAN to submit. 

## Examples

A very common task is aggregating block data to precincts.

```{r example}
library(geomander)
library(tidyverse)
 
# load precincts
data("va18sub")

# subset to target area
va18sub <- va18sub %>% filter(COUNTYFP == '087')

# create block data
block <- create_block_table(state = 'VA', county = '087')  

# match the geographies
matches <- geo_match(from = block, to = va18sub)

# Aggregate
prec <- block2prec(block_table = block, matches = matches)

```

Other important tasks include breaking data into pieces by blocks underlying them.
```{r}
library(geomander)
library(tidyverse)
 
# load precincts
data("va18sub")

# subset to target area
va18sub <- va18sub %>% filter(COUNTYFP == '087')

# create block data
block <- create_block_table(state = 'VA', county = '087')  

# Estimate down to blocks
disagg <- geo_estimate_down(from = va18sub, to = block, wts = block$VAP, value = va18sub$G18USSRSTE)

```

For more information, see the documentation and vignettes, available at [https://christopherkenny.github.io/geomander/](https://christopherkenny.github.io/geomander/)
